Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation

### 2013-10-14 exp1/present2013.tex
bug
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@671 43ef1190-90fd-4dd7-a769-d0121f02b80e
Fix the notes
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@672 43ef1190-90fd-4dd7-a769-d0121f02b80e

%\documentclass[english,aspectratio=196]]{ifislide}
%\documentclass[english,print,aspectratio=169]{ifislide}
\documentclass[english,handout,aspectratio=169]{ifislide}
\usepackage{verbatim}
\usepackage{wrapfig}
%\setbeameroption{show notes}
\usepackage[utf8]{inputenc}
% Packages used specific for this lecture:

@@

\end{frame}
\note{Natural benchmarks, query log-derived, DBPedia SPARQL Benchmark}
\note{If your queries are all fully known, by all means, run them}
\note{In most scientific relevant cases, no such assumptions can be
  made}
\note{What I will present will likely create artifical workloads}
\note{Use-case driven will seem artificial for everything else}

@@

\begin{itemize}
\item Large number of parameters to test for complex scenarios.
\item No structured approach to investigate flaws.
\item No meaningful summary of the overall performance.
\item Standardized benchmarks cannot test assertions outside of the
  standard.
\item Attempting to neutralize the effect of certain optimization
  techniques oversimplifies the test.
\end{itemize}
\end{frame}
\note{If you standardize hardware, you can't test claims about
  improved hardware. You'd need interactions too}
\note{Cache-buster is evil}

@@

\Larger 
\emph{Effects} describe the influence of the levels on the response.
\normalsize

@@

\end{frame}
\note{Interaction effects may be complex, like one implementation is
  better on CONSTRUCT queries, on small datasets}
\note{We still need to understand which effects are important and
  significant, we do that in the discussion of our experiment}

% SECTION

@@

  \end{description}
\end{frame}
\note{Implemented in R, using packages \texttt{DoE.base} and \texttt{FrF2}.}
\note{Two experiment machines, one machine running queries}
\note{No warm-up, no cache busting}
\note{256 experiments in full factorial, no replications, compare
  favorably to benchmarks}

@@

\note{Unreplicated, so we don't have a standard error, so Lenth used
  a restricted subset the median of the effects to create a Pseudo
  Standard Error}
\note{Don't worry about the axes}
\note{Yes, the Implementation is better, we knew that, joins are
  naturally strong}
\note{``Union'', ``Optional'' and ``Lang'' demanding things}
\note{``BGPComp:Lang'' and ``Lang:Union'' interactions might be due to
  that many more triples need to be searched for a language tag in one
  of the levels of the interacting factors.}
\note{But what about ``Implement::TripleC''?}
\note{We want to do a hypothesis test}
\note{Two factors do not contribute significantly to the variation
  ``Range'' and ``Machine''}

@@

\end{frame}
\note{The total variance is great, due to few runs, so few conclusions
  can be drawn}
\note{Implementation is good, and Union is hard}
\note{In the paper, we also do a hypothesis test that returns  $p =
  0.00098$. Still very low}
\note{Very rough conclusions; useful in a CI system}

@@

  \begin{itemize}
  \item Find most of the same significant effects as the full experiment.
  \item However, we see that ``Implement:TripleC'' emerges only for
  $\alpha=0.15$.
  \item ``Machine'', ``TripleC'' and ``Range'' are inactive, we can
    perform hypothesis test as above, and find $p = 5.9 \cdot 10^{-6}$.
  \end{itemize}
\end{frame}
\note{We see that ``Implement:TripleC'' emerges only for
  $\alpha=0.15$, which is important to better in real studies.}
\note{We have shown that how the experiment could be set up, we have
  shown the analysis and a hypothesis test.}
\note{Should this simplistic experiment stand up to scrutiny?}

@@

\end{frame}
\note{``Machine'' is a broad factor, good idea to group to manage
  complexity, can be broken down later}
\note{``TripleC:Machine'' and ``TripleC'' highly significant, implies
  the factors were wrongly set.}
\note{This invalidates the three preceeding experiments. We got strong
indications early on.}

@@

\begin{frame}{The role of randomization}
  \begin{itemize}
  \item We have allowed us to disregard the effect of caching,
    warm-up-runs, etc.
  \item This is different from neutralizing the effect!
  \item They are now ``lurking variables''.
  \item They contribute to the overall unexplained variance.
  \item Unexplained variance should be kept to a minimum.
  \item Time-dependent or order-dependent factors may be required.
  \end{itemize}
\end{frame}
\note{Even though we know caching is important}
\note{Caching is good practice, shouldn't be neutralized. Warm-up may
  not be permissible in real-time scenarios}
\note{Variance can obscure problems like ``Implement:TripleC''}
\note{most problematic cases are those that are
covered neither by the broad factors, nor the lurking variables, or
any of the specific factors. Invalidated by absence of language
features, such as solution modifiers.}

>>>>

%\documentclass[english,aspectratio=196]]{ifislide}
%\documentclass[english,print,aspectratio=169]{ifislide}
\documentclass[english,handout,aspectratio=169]{ifislide}
\usepackage{verbatim}
\usepackage{wrapfig}
\setbeameroption{show notes}
\usepackage[utf8]{inputenc}
% Packages used specific for this lecture:

@@

\end{frame}
\note{
  \begin{itemize}
  \item Natural benchmarks, query log-derived, DBPedia SPARQL
    Benchmark
  \item If your queries are all fully known, by all means, run them
  \item In most scientific relevant cases, no such assumptions can be
    made
  \item What I will present will likely create artifical workloads
  \item Use-case driven will seem artificial for everything else
  \end{itemize}
}

@@

\begin{itemize}
\item Large number of parameters to test for complex scenarios.
\item No structured approach to investigate flaws.
\item No meaningful summary of the overall performance.
\item Standardized benchmarks cannot test assertions outside of the
  standard.
\item Attempting to neutralize the effect of certain optimization
  techniques oversimplifies the test.
\end{itemize}
\end{frame}
\note{  
  \begin{itemize}
  \item If you standardize hardware, you can't test claims about
  improved hardware. You'd need interactions too
  \item Cache-buster is evil
\end{itemize}

}

@@

\Large
\emph{Effects} describe the influence of the levels on the response.
\normalsize

@@

\note{

  \begin{itemize}
  \item Interaction effects may be complex, like one implementation is
    better on CONSTRUCT queries, on small datasets
\item We still need to understand which effects are important and
  significant, we do that in the discussion of our experiment
\end{itemize}
}
% SECTION

@@

  \end{description}
\end{frame}

\note{

  \begin{itemize}
  \item Implemented in R, using packages \texttt{DoE.base} and
    \texttt{FrF2}.
  \item Two experiment machines, one machine running queries
  \item No warm-up, no cache busting
  \item 256 experiments in full factorial, no replications, compare
    favorably to benchmarks
  \end{itemize}
}

@@

\note{
  \begin{itemize}
  \item Unreplicated, so we don't have a standard error, so Lenth used
    a restricted subset the median of the effects to create a Pseudo
    Standard Error
  \item Don't worry about the axes
  \item Yes, the Implementation is better, we knew that, joins are
    naturally strong
  \item ``Union'', ``Optional'' and ``Lang'' demanding things
  \item ``BGPComp:Lang'' and ``Lang:Union'' interactions might be due
    to that many more triples need to be searched for a language tag
    in one of the levels of the interacting factors.
  \item But what about ``Implement::TripleC''?
  \item We want to do a hypothesis test
  \item Two factors do not contribute significantly to the variation
    ``Range'' and ``Machine''
  \end{itemize}
}

@@

\end{frame}
\note{

  \begin{itemize}
  \item The total variance is great, due to few runs, so few
    conclusions can be drawn
  \item Implementation is good, and Union is hard
  \item In the paper, we also do a hypothesis test that returns $p =
    0.00098$. Still very low
  \item Very rough conclusions; useful in a CI system
  \end{itemize}
}

@@

  \begin{itemize}
  \item Find most of the same significant effects as the full experiment.
  \item However, we see that ``Implement:TripleC'' emerges only for
  $\alpha=0.15$.
  \item ``Machine'', ``TripleC'' and ``Range'' are inactive, we can
    perform hypothesis test as above, and find $p = 5.9 \cdot 10^{-6}$.
  \end{itemize}
\end{frame}
\note{
  \begin{itemize}
  \item We see that ``Implement:TripleC'' emerges only for
    $\alpha=0.15$, which is important to better in real studies.
  \item We have shown that how the experiment could be set up, we have
    shown the analysis and a hypothesis test.
  \item Should this simplistic experiment stand up to scrutiny?
  \end{itemize}
}

@@

\end{frame}
\note{
  \begin{itemize}
  \item ``Machine'' is a broad factor, good idea to group to manage
    complexity, can be broken down later
  \item ``TripleC:Machine'' and ``TripleC'' highly significant,
    implies the factors were wrongly set.
  \item This invalidates the three preceeding experiments. We got
    strong indications early on.
  \end{itemize}
}

@@

\begin{frame}{The role of randomization}
  \begin{itemize}
  \item We have allowed us to disregard the effect of caching,
    warm-up-runs, etc.
  \item This is different from neutralizing the effect!
  \item They are now ``lurking variables''.
  \item They contribute to the overall unexplained variance.
  \item Unexplained variance should be kept to a minimum.
  \item Time-dependent or order-dependent factors may be required.
  \end{itemize}
\end{frame}
\note{
  \begin{itemize}
  \item Even though we know caching is important
  \item Caching is good practice, shouldn't be neutralized. Warm-up
    may not be permissible in real-time scenarios
  \item Variance can obscure problems like ``Implement:TripleC''
  \item most problematic cases are those that are covered neither by
    the broad factors, nor the lurking variables, or any of the
    specific factors. Invalidated by absence of language features,
    such as solution modifiers.
  \end{itemize}
}
